---
title: "code_Tian"
author: "Tian Walker"
date: "2023-03-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
#install.packages("plotly")
library(plotly)
library(rio)
library(here)
```

```{r}
data <- import(here("reg_data.csv"))
dat <- import(here("pre_reg_data.csv"))
```

Hello Cengiz! 

My goal is to move this graph into plotly and have it show which resources people used when you hover on the dot. So the x axis tells you how many resources a person used and then when you hover you find out which resources. 

The data is transformed from the name of the resource to a 0 or 1 in order to create the variable that shows the total number of resources used. 
The dataset called "dat" above has the data in its raw form. I have recreated the steps to transform the variable and then sum the rows below. 
I am thinking that I would need to duplicate all of the variables before I transform them so that the data retains them in their raw form as well. 

Then, the real issue is whether or not plotly will show a different variable when you hover than what is on the x or y axis. 

Thank you for your time! 

-Tian

```{r}
data %>% 
  ggplot(aes(res_sum, des_score)) + 
  geom_point(aes(size = chai_means),color = "purple3") +           scale_color_manual(values = my_colors)+ 
  theme_classic() + 
  theme(axis.line = element_line(colour = "darkmagenta", 
       linewidth = 2, linetype = "solid")) +  scale_fill_manual(values = my_colors) + 
  xlim("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13") +
  labs(x = "Number of Resources Used", y = "Diabetes Empowerment", size = "Health Engagement Score")
```

```{r}

dat %>% 
  select(starts_with("resource"))

```
```{r}
dat %>% 
   mutate(
    across(
      # ^means begins with. 0-9 are the possible numbers and {1,2} means it 
      # could be one or two digits
      matches("^resources_[0-9]{1,2}$"), 
      # need to ask Sara about what the period inside parens indicates
      ~ ifelse(is.na(.), 0, 1)))

```
```{r}
data$res_sum <- data %>% 
  select(resources_1:resources_21)%>%
  rowSums(na.rm=TRUE)
```

```{r}
data %>% 
  ggplot(aes(res_sum, des_score)) + 
  geom_point(aes(size = chai_means),color = "purple3") +           scale_color_manual(values = my_colors)+ 
  theme_classic() + 
  theme(axis.line = element_line(colour = "darkmagenta", 
       linewidth = 2, linetype = "solid")) +  scale_fill_manual(values = my_colors) + 
  xlim("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13") +
  labs(x = "Number of Resources Used", y = "Diabetes Empowerment", size = "Health Engagement Score")
```

