---
title: "Lab 1"
author: "Student Name"
date: "YYYY-MM-DD"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data

We'll work with the [#tidytuesday](https://twitter.com/search?q=%23tidytuesday&src=tyah)  data for 2019, specifically the [#rstats](https://twitter.com/search?q=%23rstats&src=typd) dataset, containing nearly 500,000 tweets over a little more than a decade using that hashtag. 

The data is in under Dataset tab of Week 3 module on Canvas.

You can import the dataset using the code below.

```{r}
library(tidyverse)
library(tidytext)
```


```{r include=FALSE}
d <- rio::import(here::here("lab1", "rstats_tweets.rds"),
                 setclass = "tbl_df") 

#View(d)
```

If you need help with processing text data, please revisit the notebook introduced in Week 1.

https://www.kaggle.com/code/uocoeeds/introduction-to-textual-data

### Histogram and Density plots

1. Create a histogram the column `display_text_width` using the `ggplot2` package and `geom_histogram()` function. Try at least four different numbers of bins (e.g., 20, 30, 40, 50) by manipulating the `bins=` argument. Select what you think best represents the data for each. Provide a brief justification for your decision. For all plots you created, change the default background color from grayish to white.

```{r, echo=F, eval=F}

d %>% 
  select(display_text_width) %>% 
  view()

d %>% 
  ggplot(aes(x = display_text_width))+ geom_histogram() + theme_bw()
         

# p <- ggplot(d, aes(x=display_text_width)) + 
#   geom_histogram()

```

```{r}
d %>% 
  ggplot(aes(x = display_text_width))+ geom_histogram(bins = 20, color = "black") + theme_bw()

d %>% 
  ggplot(aes(x = display_text_width))+ geom_histogram(bins = 30, color = "black") + theme_bw()

d %>% 
  ggplot(aes(x = display_text_width))+ geom_histogram(bins = 40, color = "black") + theme_bw()
         
d %>% 
  ggplot(aes(x = display_text_width))+ geom_histogram(bins = 50, color = "black") + theme_bw()
```

2. Create a density plot for the column `display_text_width` using the `ggplot2` package and `geom_density()` function. Fill the inside of density plot with a color using the `fill=` argument. Try at least four different numbers of smoothing badwith (e.g., 0.2, 1.5, 3, 5) by manipulating the `bw=` argument. Select what you think best represents the data for each. Provide a brief justification for your decision.


```{r, echo=TRUE, eval=TRUE}
d %>% 
  ggplot(aes(x = display_text_width))+ geom_density(fill = "deeppink4") + theme_bw()
```

```{r eval=FALSE, include=FALSE}
library(stringr)

a <- LETTERS
remove_words<- c('Z', 'G', 'K', 'M', 'S')

str_extract(string = a, 
            pattern = remove_words)

str_extract(string = a, 
            pattern = 'K', 
            simplify = TRUE)
```

```{r eval=FALSE, include=FALSE}
dplyr::filter(.)

a <- data.frame(letters = LETTERS)

grepl
```


```{r eval=FALSE, include=FALSE}

?setdiff
d <- setdiff( x = a$letters, 
         y = remove_words)

a %>% 
  setdiff()

a_reduced <- a[!a$letters %in% remove_words,]

```

### Barplot

3. Using the information `text` column, create the following figure of the 15 most common words represented in these posts by using the `ggplot2()` package and `geom_col()` function. Remove the stop words, and also exclude the words such as 't.co','https','http','rt','rstats'.

```{r eval=FALSE, include=FALSE}


View(d)

d$text %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = reorder(word, n)) %>% # make y-axis ordered by n
  slice(1:15) %>% # select only the first 15 rows
  ggplot(aes(n, word)) +
    geom_col(fill = "cornflowerblue")


d <- tibble(
  paragraph = seq_along(text),
  description = text
)

text

view(d)
```

```{r}

# eugene_df <- tibble(
#   paragraph = seq_along(d$text),
#   description = d$text
# )
# 
# ?seq_along
# 
# eugene_df

names(d)
dat <- d %>% 
  select(user_id, text)

dat <- dat %>% 
unnest_tokens(word, text)

dat
```

```{r}
#dat_reduced <- dat[!dat$text %in% stop_words,]


dat %>% 
anti_join(stop_words) %>% 
filter(word != "t.co",word !="https",word != "http",word != "rt", word !="rstats") %>% 
count(word, sort = TRUE) %>% 
  mutate(word = reorder(word, n)) %>% # make y-axis ordered by n
  slice(1:15) %>% # select only the first 15 rows
  ggplot(aes(n, word)) +
  theme_minimal() +
    geom_col(fill = "cornflowerblue")+ theme(panel.grid.major.x = element_line(color = "grey",size = 0.25, linetype = 1), panel.grid = element_blank()) + labs( x = "Count", y = "Word", Title = "Word frequencies in posts", subtitle = "Top 15 wordss displayed", caption = "Data from Mike Kearny, distributed from #tidytuesday")
                                          
  
  
  
  #theme(panel.grid.major = element_line(color = "black",
                                        # size = 0.5,
                                        # linetype = 1)) 
```



```{r,echo=TRUE,eval=TRUE}

```



4. Style the plot so it (mostly) matches the below. It does not need to be exact, but it should be close.

```{r, echo=TRUE,eval=TRUE}

```

